The \kplr\ Catalog: Systematic probabilistic parameter estimation for every \kepler\ planet candidate

### 2013-12-12 document/ms.tex
document: inferring the empirical radius distribution

>>>>

\section{The empirical radius distribution}

The problem of determining the empirical distribution of exoplanet radii
without taking selection effects into account (see Foreman-Mackey \etal\ in
prep), can be written as a hierarchical inference problem.
We have samples from the posterior probability distribution
\begin{eqnarray}
z^{(j)} &\sim& p(z\,|\,x,\,\alpha_0)
\end{eqnarray}
where $z$ is the radius ratio, $x$ is the light curve, and $\alpha_0$ are the
parameters of the interim prior $p(z\,|\,\alpha_0)$.
Instead, we would like to compute and/or sample the parameters describing the
distribution of physical radii $p(r\,|\,\alpha)$.
To do this, we must compute the \emph{marginalized likelihood}
\begin{eqnarray}\eqlabel{marg-like}
p(x,\,\hat{R}\,|\,\alpha) &=&
    \int \dd r\dd R\dd z\,\prod_n p(R_n)\,p(\hat{R}_n\,|\,R_n)
        \prod_k p(r_{nk}\,|\,\alpha)\,p(z_{nk}\,|\,R_n,\,r_{nk})\,
                p(x\,|\,z_{nk}) \quad.
\end{eqnarray}
We can rewrite the inner term---using the fact that
$p(z_{nk}\,|\,R_n,\,r_{nk})\equiv\delta(r_{nk}-R_n\cdot z_{nk})$---as
\begin{eqnarray}
p_{nk}(x\,|\,R_n,\,\alpha)&=&
\int \dd r_{nk}\dd z_{nk}\,p(r_{nk}\,|\,\alpha)\,p(z_{nk}\,|\,R_n,\,r_{nk})\,
        p(x\,|\,z_{nk}) \nonumber\\
&=&
\int \dd z_{nk}\,p(z_{nk} \cdot R_n\,|\,\alpha)\,p(x\,|\,z_{nk}) \nonumber\\
&=&
\int \dd z_{nk}\,p(z_{nk}\cdot R_n\,|\,\alpha)\,p(x\,|\,z_{nk})\,
\frac{p(z_{nk}\,|\,x,\,\alpha_0)}{p(z_{nk}\,|\,x,\,\alpha_0)} \nonumber\\
&\propto&
\int \dd z_{nk}\,\frac{p(z_{nk}\cdot R_n\,|\,\alpha)}{p(z_{nk}\,|\,\alpha_0)}\,
p(z_{nk}\,|\,x,\,\alpha_0) \nonumber\\
&\approx&
\frac{1}{J_{nk}}
\sum_j \frac{p(z_{nk}^{(j)}\cdot R_n\,|\,\alpha)}{p(z_{nk}^{(j)}\,|\,\alpha_0)}
\end{eqnarray}

The integral over stellar radius can also be approximated using posterior
samples
\begin{eqnarray}
R^{(l)} &\sim& p(R\,|\,\hat{R}) \quad.
\end{eqnarray}
Here, we're not trying to simultaneously infer the distribution of stellar
parameters.
This simplifies matters and the full integral in \eq{marg-like} can be
approximated as
\begin{eqnarray}
\frac{p(x,\,\hat{R}\,|\,\alpha)}{p(x,\,\hat{R}\,|\,\alpha_0)}
&\approx& \prod_n \frac{1}{L_n} \sum_{l=1}^{L_n} \prod_k
\frac{1}{J_{nk}} \sum_{j=1}^{J_{nk}}
\frac{p(z_{nk}^{(j)}\cdot R_n\,|\,\alpha)}{p(z_{nk}^{(j)}\,|\,\alpha_0)}
\quad.
\end{eqnarray}
This ratio can now be computed efficiently for different values $\alpha$ and
optimized or you can set priors $p(\alpha)$ and sample.
